Nanocall: An Open Source Basecaller for Oxford Nanopore Sequencing Data

### 2016-03-17 paper-biorxiv/paper.tex
sequencing overview edits

\paragraph{Sequencing overview.}
Informally, the MinION sequencing works as follows. As part of an initial library preparation step, DNA is sheared into fragments of 8--20 Kbp. Proprietary chemistry guides the resulting double-stranded DNA fragments through a grid of small-sized holes, called ``nanopores'', where the actual sequencing takes place. To that end, adapters are ligated and the complementary strands of each DNA fragment are separated, resulting in a single-stranded DNA product that contains the original two strands concatenated, one after the other. (In ONT terminology, the first strand going through the nanopore is the \emph{template}, and the second one is the \emph{complement}.) The single-stranded DNA product is threaded through a nanopore, one DNA base at a time, all while measuring an electrical current that traverses the nanopore. The central idea is that the single-stranded DNA product present in the nanopore affects the current in a way that is strong enough to enable an eventual decoding process. This process, called \emph{basecalling}, takes as input a list of current measurements, and produces as output a list of DNA bases most likely to have generated those currents.

>>>>

\paragraph{Sequencing overview.}
Informally, the MinION sequencer works as follows. First, DNA is sheared into fragments of 8--20 Kbp and adapters are ligated to either end of the fragments. The resulting DNA fragments pass through a protein embedded in a membrane via a nanometre-sized channel (this protein is the ``nanopore''). A single strand of DNA passes through the pore; the optional use of a hairpin adapter at one end of the fragment allows the two strands of DNA to serially pass through nanopore, allowing two measurements of the fragment. In ONT terminology, the first strand going through the nanopore is the \emph{template}, and the second is the \emph{complement}. As a DNA strand passes through the pore it partially blocks the flow of electric current through the pore. The flow of current is sampled over time which is the observable output of the system. The central idea is that the single-stranded DNA product present in the nanopore affects the current in a way that is strong enough to enable decoding the electric signal data into a DNA sequence. This process, called \emph{basecalling}, takes as input a list of current measurements, and produces as output a list of DNA bases most likely to have generated those currents.
